Type Ia Supernova Explosions from Hybrid Carbon-Oxygen-Neon White Dwarf Progenitors

### 2019-02-11 paper/ms.tex
Finished the methodology section of Carlyn's paper. On to discussion of the results.

Within this progenitor setting, many mechanisms for the explosion have 
been studied with the result being that the outcome is sensitive to
the ignition conditions and the nature of the thermonuclear buring.
Studies indicate that neither a pure deflagration (subsonic flame front)
nor a pure detonation (supersonic burning front) produce ejecta conisitent
with observations \citep{arnett69,roepkeetal07}.  Instead, the models 
that best reproduce the observed stratified ejecta are those in which 
the burning begins as a subsonic deflagration, which allows the star 
to react and expand thus lowering the density, and is then followed by a 
supersonic detonation that incinerates the 
star~\citep{Nomo84,Khokhlov1991Delayed-detonat,HoefKhok96,GameKhokOran05}.
The ``classic'' delayed detonation model is that of 
\citet{Khokhlov1991Delayed-detonat} \citep[See also][]{hoflich.khokhlov.ea:delayed,GameKhokOran05},
and many variations have been 
explored \citep[and references therein]{hillebrandtetal2013,calderetal2013}.
The approach we employ for this study is a variation, the 
deflagration-to-detonation transition (DDT), in which the transition
occurs at the top of a rising Rayleigh-Taylor unstable plume of hot, burned 
material when it reaches a threshold density.
We describe this approach and our methodology in detail below.

@@

The mechanism by which the deflagration transistions to a detonation
is imcompletely understood, but the development of fluid instabilities
at the deflagration front, particuarly the Rayleigh-Taylor 
instability~\citep{taylor+50,chandra+81}, is an accepted part of the process.
Many ideas have been proposed, including the deflagration front
entering a regime of distributed burning and the net
rate becoming supersonic \citep{NiemWoos97}, the unstable
flame reaching a certian fractal dimension~\citep{woosley90},
and the Zel'dovich mechanism in which condtions are right for
a pocket of fuel to 
detonate~\citep{zeldovichetal1970,KhokOranWhee97,jacketal2014}.
We note that regardless of
the mechanism, the amount of expansion of the star prior to the bulk of burning
is critical to the composition of material synthesized in the explosion. Also,
early burning during the deflagtation phase is at high enough densites that
the effects of electron capture are significant and similarly influence the 
compositon the material synthesised in the explosion~\citep{hoeflichetal2004,
hoeflich2006,fesenetal2007,diamondetal2018}.
Our simulations assume that the transition occurs when the top of a 
rising, Rayleigh-Taylor unstable plume reaches a characteristic low
density~\citep{townsley.calder.ea:flame}. 
Variations on the choice of this density allowed earlier studies to explore
the effects of physics influencing the transition 
density~\citep{jacketal2010,Krueger2010On-Variations-o,kruegetal12}.

@@

From population synthesis, \citet{mengpods2014} found that these
%meng received may 5 accepted may 16
progenitors may substantially contribue to the population of SNeIa (1-8\%) and have
relatively short delay times. They also suggested that these 
may produce part of the Iax class of events. \citet{Wangetal2014} also with population
synthesis studied the case
%received august 28 accepted sept 26
of a hybrid progenitor accreting from a nondegenerate He star and found
birth rates indicating that up to 18\% of SNe Ia may follow from this chanel
and very short delay times. \citet{Wangetal2014} also suggested that explosions 
from hybrid progenitors may provide an explanation for type Iax events.
\citet{mengpods2018}, from the common-envelope-wind model developed in 
\citep{mengpods2014}, propose that both Ia-CSM and Iax events 
are caused by the explosion of hybrid progenitors, with Ia-CSM occuring in systems with 
a massive common envelope and Iax events occuring in systems where most of the common envelope
has been lost.

@@

The reaction progress variable is $\phi$ and it varies from $\phi=0$ for
unburned fuel and $\phi=1$ for burned ash. $\phi$ is evolved
via an advection-diffusion-reaction equation,
\begin{equation}
  \label{eq:ard}
  \partial_t \pv + \vec{u}\cdot\nabla \pv = \kappa \nabla^2 \pv +
\frac{1}{\tau} R\left(\phi\right) ,
\end{equation}
where $\vec{u}$ is the velocity of the fluid, $\kappa$ is the
diffusion coefficient, $\tau$ is the reaction timescale, and $R(\phi)$ is
a non-dimensional function describing the reaction. The parameters
$\kappa$, $\tau$, and $R(\phi)$ are tuned to propagate the reaction
front at a prescribed speed.  Our model uses the ``sharpened KPP
 described by~\cite{VladWeirRyzh06}, 
with $R\propto(\phi-\epsilon)(1-\phi+\epsilon)$, where
$\epsilon \simeq 10^{-3}$.  This scheme has been shown to be
acoustically quiet, stable, and to give a unique flame speed
\cite{townsley.calder.ea:flame}. The input flame speeds come from
tabulated results obtained by direct numerical simulations of
thermonuclear burning~\cite{timmes92,Chamulak2007The-Laminar-Fla},
these are boosted to account for speed-up of burning due to unresolved 
buoyancy and background turbulence~\cite{Khok95,
gamezo.khokhlov.ea:thermonuclear,townsley.calder.ea:flame,jacketal2014}.

@@

\subsection{Ttwo-dimensional Models from One-dimensional Models}

The initial progenitor models for the two-dimensional simulations were
created from the one-dimensional \MESA\ models by mapping the one-dimensional
models onto the two-dimensional domain while 
preserving hydrostatic equilibrium.
As described above, the flame capturing scheme with progress
variables describing the evolution as the material burns. The \MESA\
models, however, relied on a detailed reaction network that included
many species. Accordingly, the process of creating the two-dimensional
models for \FLASH\ required aggregating some of the species abundances.
As with much of this study, we applied the 
techniques of ~\citet{willcoxetal2016}.

The process of mapping the \MESA\ models began by converting 
the \MESA\ model to a uniform grid of $4$~km resolution. \MESA\ 
quantities in zones with spacing less than $4$~km 
were mass-weighted averaged and a quadratic interpolation
was used to calculate quantities for the case of \MESA\ zones having
a spacing greater than $4$~km. At this point, 
abundances of nuclides from the \MESA\ models were aggregated
into the abundances tracked by the flame capturing scheme in
\FLASH.
The most abundant isotopes in the model were \C{12}, \Ox{16},
and \Ne{20}, which are symmetric (the number of neutrons 
equals the number of protons). Similarly to~\citet{willcoxetal2016}, 
other neutron-rich isotopes in the initial model were combined into 
\Ne{22}, which serves as a proxy for metallicty. The aggregation
account for the $Y_e$ of the full set of nuclides, \Ne{20} 
and \Ox{16} were constrained to be in the same ratio in both 
sets of abundances.

@@

The fuel starts off cold. The elections in the plasma conduct heat and the
fuel begins to heat up. Once the fuel reaches a certain temperature, the
reaction starts. The fuel abundance begins to decline because it is consuming
fuel. Energy is produced which makes the fuel hot. Eventually, all that’s
left is ash. As surface area grows, conducted energy increases and density
decreases. At low densities Rayleigh-Taylor instability causes the flame to
get more tangled and unstable. Eventually, the star will get a large enough
surface area in this volume that the net burning effect is supersonic.
Therefore, the transition from deflagration to detonation is made.

Using the FLASH code from University of Chicago, simulations of
explosions were preformed. For each simulation we varied the central
ignition seed.

----updated text-----

DDT Process:
The fuel starts off cold. The elections in the plasma conduct heat
and the fuel begins to heat up. Once the fuel reaches a certain
temperature, the reaction starts. The fuel abundance begins to
decline because it is consuming fuel. Energy is produced which
makes the fuel hot. Eventually, all that’s left is ash. As
surface area grows, conducted energy increases and density
decreases. At low densities Rayleigh-Taylor instability causes
the flame to get more tangled and unstable. Eventually, the
star will get a large enough surface area in this volume that
the net burning effect is supersonic. Therefore, the transition
from deflagration to detonation is made.

A deflagration alone will not produce an event of normal
brightness and expansion velocity. Instead, the initial
deflagration must transition to a detonation after the star has
expanded some in order to produce abundances and a stratified
ejecta in keeping with observations. The physics of this
“deflagration-to- detonation transition” (DDT) are not
completely understood, but there has been considerable study
based on mechanisms involving flame fronts in highly turbulent
conditions. These models generally reproduce the observations
under certain assumptions about the ignition, but research has
shown that the results are very sensitive to the details of the
ignition. In our simulations, we initialize a detonation once
the deflagration front reaches a characteristic DDT fuel density,
which controls the degree of expansion the star undergoes during
the deflagration stage.

The comparison study of tradition C/O progenitor WDs used the same initial
conditions for the realizations within the suite. The traditional models
are simulated with the same code and the same set of perturbations on the
initial match head (realization). The only difference is the different
nuclear burning to include the other Ne and the initial profile of the
1-d model.

----above updated text-----

>>>>

Within this progenitor setting, many mechanisms for the explosion have 
been studied with the result being that the outcome is sensitive to
the ignition conditions and the nature of the thermonuclear buring.
Studies indicate that neither a pure deflagration (subsonic flame front)
nor a pure detonation (supersonic burning front) produce ejecta conisitent
with observations \citep{arnett69,roepkeetal07}.  Instead, the models 
that best reproduce the observed stratified ejecta are those in which 
the burning begins as a subsonic deflagration, which allows the star 
to react and expand thus lowering the density, and is then followed by a 
supersonic detonation that incinerates the 
star~\citep{Nomo84,Khokhlov1991Delayed-detonat,HoefKhok96,GameKhokOran05}.
The ``classic'' delayed detonation model is that of 
\citet{Khokhlov1991Delayed-detonat} \citep[See also][]{hoflich.khokhlov.ea:delayed,GameKhokOran05},
and many variations have been 
explored \citep[and references therein]{hillebrandtetal2013,calderetal2013}.
The approach we employ for this study is a variation, the 
deflagration-to-detonation transition (DDT), in which the transition
occurs at the top of a rising Rayleigh-Taylor~\citep{taylor+50,chandra+81} 
unstable plume of hot, burned 
material when it reaches a threshold density.
We describe this approach and our methodology in detail below.

@@

The mechanism by which the deflagration transistions to a detonation
is imcompletely understood, but the development of fluid instabilities
at the deflagration front, particuarly the Rayleigh-Taylor 
instability, is an accepted part of the process.
Many ideas have been proposed, including the deflagration front
entering a regime of distributed burning and the net
rate becoming supersonic \citep{NiemWoos97}, the unstable
flame reaching a certian fractal dimension~\citep{woosley90},
and the Zel'dovich mechanism in which condtions are right for
a pocket of fuel to 
detonate~\citep{zeldovichetal1970,KhokOranWhee97,jacketal2014}.
We note that regardless of
the mechanism, the amount of expansion of the star prior to the bulk of burning
is critical to the composition of material synthesized in the explosion. Also,
early burning during the deflagtation phase is at high enough densites that
the effects of electron capture are significant and similarly influence the 
compositon the material synthesised in the explosion~\citep{hoeflichetal2004,
hoeflich2006,fesenetal2007,diamondetal2018}.
Our simulations assume that the transition occurs when the top of a 
rising, Rayleigh-Taylor unstable plume reaches a characteristic low
density~\citep{townsley.calder.ea:flame}. 
Variations on the choice of this density allowed earlier studies to explore
the effects of physics influencing the transition 
density~\citep{jacketal2010,Krueger2010On-Variations-o,kruegetal12}.

@@

From population synthesis, \citet{mengpods2014} found that these
%meng received may 5 accepted may 16
progenitors may substantially contribue to the population of \SNeIa (1-8\%) and have
relatively short delay times. They also suggested that these 
may produce part of the Iax class of events. \citet{Wangetal2014} also with population
synthesis studied the case
%received august 28 accepted sept 26
of a hybrid progenitor accreting from a nondegenerate He star and found
birth rates indicating that up to 18\% of SNe Ia may follow from this chanel
and very short delay times. \citet{Wangetal2014} also suggested that explosions 
from hybrid progenitors may provide an explanation for type Iax events.
\citet{mengpods2018}, from the common-envelope-wind model developed in 
\citep{mengpods2014}, propose that both Ia-CSM and Iax events 
are caused by the explosion of hybrid progenitors, with Ia-CSM occuring in systems with 
a massive common envelope and Iax events occuring in systems where most of the common envelope
has been lost.

@@

The reaction progress variable is $\phi$ and it varies from $\phi=0$ for
unburned fuel and $\phi=1$ for burned ash. $\phi$ is evolved
via an advection-diffusion-reaction equation,
\begin{equation}
  \label{eq:ard}
  \partial_t \pv + \vec{u}\cdot\nabla \pv = \kappa \nabla^2 \pv +
\frac{1}{\tau} R\left(\phi\right) ,
\end{equation}
where $\vec{u}$ is the velocity of the fluid, $\kappa$ is the
diffusion coefficient, $\tau$ is the reaction timescale, and $R(\phi)$ is
a non-dimensional function describing the reaction. The parameters
$\kappa$, $\tau$, and $R(\phi)$ are tuned to propagate the reaction
front at a prescribed speed.  Our model uses the ``sharpened KPP
 described by~\cite{VladWeirRyzh06}, 
with $R\propto(\phi-\epsilon)(1-\phi+\epsilon)$, where
$\epsilon \simeq 10^{-3}$.  This scheme has been shown to be
acoustically quiet, stable, and to give a unique flame speed
\cite{townsley.calder.ea:flame}. The input flame speeds come from
tabulated results obtained by direct numerical simulations of
thermonuclear burning~\cite{timmes92,Chamulak2007The-Laminar-Fla},
and 
these are boosted to account for speed-up of burning due to unresolved 
buoyancy and background turbulence~\cite{Khok95,
gamezo.khokhlov.ea:thermonuclear,townsley.calder.ea:flame,jacketal2014}.

The two-dimensional models in this study do not utilize a subgrid-scale
model for the turbulence-flame 
interaction~\citep[See][for examples]{Schmetal06a,Schmetal06b,jacketal2014},
The models only apply in three-dimensional
simulations beacuse two-dimensional hydrodynamics cannot
correctly describe turbulence. The simulations we present
use the mimimal enhancement based on the Rayleigh-Taylor
strength introduced by~\citet{townsley.calder.ea:flame}. The
assumption here is that the burning self-regulate on resolved 
scales so that results are insensitive to the detailed treatment 
of the interaction with turbulence, and previous experience
indicates this assumption is reasonable for comparisons
like the one presented in this work~\citet{townsley.calder.ea:flame,
willcoxetal2016}.

@@

\subsection{Two-dimensional Initial Conditions From One-dimensional Models}

The initial progenitor models for the two-dimensional simulations were
created from the one-dimensional \MESA\ models by mapping the one-dimensional
models onto the two-dimensional domain while 
preserving hydrostatic equilibrium.
As described above, the flame capturing scheme with progress
variables describing the evolution as the material burns. The \MESA\
models, however, relied on a detailed reaction network that included
many species. Accordingly, the process of creating the two-dimensional
models for \FLASH\ required aggregating some of the species abundances.
As with much of this study, we applied the 
techniques of~\citet{willcoxetal2016}.

The process of mapping the \MESA\ models began by converting 
the \MESA\ model to a uniform grid of $4$~km resolution. \MESA\ 
quantities in zones with spacing less than $4$~km 
were mass-weighted averaged and a quadratic interpolation
was used to calculate quantities for the case of \MESA\ zones having
a spacing greater than $4$~km. At this point, 
abundances of nuclides from the \MESA\ models were aggregated
into the abundances tracked by the flame capturing scheme in
\FLASH.
The most abundant isotopes in the model were \C{12}, \Ox{16},
and \Ne{20}, which are symmetric (the number of neutrons 
equals the number of protons). Similarly to~\citet{willcoxetal2016}, 
other neutron-rich isotopes in the initial model were combined into 
\Ne{22}, which serves as a proxy for metallicty. The aggregation
account for the $Y_e$ of the full set of nuclides, \Ne{20} 
and \Ox{16} were constrained to be in the same ratio in both 
sets of abundances. \figref{fig:init_conds} shows the initial
profile of the one-dimensional models.

@@

The simulations performed for this study consisted to a suite of
{\color{red} 35 CHECK THIS! 40?} two-dimensional simulations of 
thermonuclear supernova
explosions from hybrid C-O-Ne progenitors in the DDT explosion paradigm. These 
were compared to a suite of supernova simulations from traditional
C-O progenitors. In both suites, the simulation
begins with a progenitor model mapped to the two-dimensional 
\FLASH\ grid. The burning is initiated with a ``matchhead,'' a region 
in the white dwarf's center that is fully burned to nuclear statistical
equilibrium (NSE). 
This initially burned regeion ignites a deflagration, a subsonic
flame, and because the mathhead was perturbed it is unstable to
the Rayleigh-Taylor instability with the result that buoyant plumes 
rise. The star is partially consumed during this deflagration phase,
and the star responds by expanding.
When a plume reaches a specified density,
a detonation is initiated, and the simulation continues until 
the expanding star reaches low densities at which point burning
effectively ends. The paragraphs in this section provide details
of the implementation of this method.

{\color{red} ARE THE BELOW NUMBERS CORRECT?}
The C-O and C-O-Ne models have the same central density
and temperature $7 \times 10^8$~K, respectively.
The C-O model had core composition of (\C{12} = 0.4, 
\Ne{20} = 0.03, \Ox{16} = 0.57) and envelope composition 
of (\C{12} = 0.5, \Ne{20} = 0.02, \Ox{16} = 0.48). 
The C-O-Ne model had core composition of (\C{12} = 0.X, 
\Ne{20} = 0.XX, \Ox{16} = 0.XX) and envelope composition 
of (\C{12} = 0.XX, \Ne{20} = 0.0X, \Ox{16} = 0.XX). 
\figref{fig:co_vs_cone_temp} show the density and temperature 
profiles of the C-O and C-O-Ne hybrid WD.

The simulations were performed in two-dimensional $\mathrm{z-r}$
cylindrical coordinates, extending radially from 0 to
$65,536$~\km\ and along the axis of symmetry from $-65,536$~\km\ to
$65,536$~\km. The maximimum refinement level of the adaptive mesh 
corresponded to $4$~\km\ resolution, which previous study has shown is
is a good balance between efficiency and 
accuracy~\citep{townsley.calder.ea:flame,townetal2009}.
This resolution and geometry was used in previous studies
allowing direct comparison to previous results~\citep{kruegetal12}.

The ignition of the deflagration via a matchhead followed the 
initialization described in~\citet{kruegetal12}, which
followed the method of~\citet{townetal2009}.
In both the hybrid and traditional cases, the matchhead had a 
nominal radius of $150$~km before 
a different randomly-seeded perturbation was applied to the 
matchhead for each of the 35 simulations in both suites. {\color{red}
IS THIS RIGHT- BOTH USED THE SAME REALIZATIONS?}
The perturbation to the sphere's surface is a set of
spherical harmonic functions with randomly chosen amplitudes, and
each set of perturbations is referred to as a ``realization." 
These perturbations have been shown to reproduce 
the scatter in \Ni{56} 
yield from \SNeIa\ \citet{townetal2009}.

The transition from deflagration to detonation again follows
the previous studies, which assumed the transition location
is parameterized by the fuel density $\rho_{\mathrm{DDT}}$.
When a rising
plume reaches the threshold density, in this case
$\rho_{\mathrm{DDT}} = 10^{7.2}~\grampercc$, 
a $12$~\km\ a region of fuel, (\C{12} and \Ne{20}, if present) is
fully burned $32$~\km\ radially outwards from the flame. This
instantaneous burning in the region of this size provides conditions
to generate the shock and support the detonation at the chosen
threshold density.  Multiple DDT points may arise, but they are
constrained to be at least $200$~\km\ apart. The choice of DDT
in the suite has been shown to be high enough to ensure 
the robust ignition of a detonation shock.

Once the detonation starts, the remaining fuel at densities
high enough for the detonation to propagate is quickly consumed.
The simulations were run to 4.0 s, at which time burning 
has effectively ceased.
